% arara: pdflatex
% arara: biber
% arara: pdflatex
% arara: pdflatex
\documentclass[
  tufte-book,
  notoc,
  symmetric,
  biblatex={
    citestyle=authoryear-comp,
    % citestyle=authortitle-icomp,
    autocite=footnote,
    maxcitenames=2,
    bibstyle=authoryear,
    dashed=false,
    mergedate=basic,
    maxbibnames=99,
    backref=true,
    doi=false,
    url=false,
    isbn=false,
  }
]{tufte-thesis}

\usepackage{xpatch}
% \xapptobibmacro{cite}{\setunit{\nametitledelim}\printfield{year}}{}{}

\input{preamble}

\geometry{letterpaper,left=1.25in,top=1in,headsep=2\baselineskip,textwidth=26pc,marginparsep=2pc,marginparwidth=12pc,textheight=44\baselineskip,headheight=\baselineskip}


\usepackage{cmu-titlepage2}


\addbibresource{thesis.bib}

%% Workaround to \NoCaseChange and \tl_map_tokens:nn issue:
% \ExplSyntaxOn
% \cs_gset:Npn \NoCaseChange #1 { \use:n {#1} }
% \ExplSyntaxOff

\begin{document}

\frontmatter
% \pagenumbering{roman}
% \pagestyle{empty}

\title{Session-Typed Ordered Logical Specifications}

\author{Henry DeYoung}
\date{December 2020}
\Year{2020}
\trnumber{CMU-CS-20-133}

\committee{%
  Frank Pfenning (Chair)\\
  Iliano Cervesato\\
  Robert Harper\\
  Andr\'{e} Platzer\\
  Simon Gay (University of Glasgow)\\
  Carsten Sch\"{u}rmann (IT University of Copenhagen)%
}

\support{This research was sponsored by National Science Foundation award number: 1718276; by a National Science Foundation Graduate Research Fellowship award; and by a Google Lime Scholarship award.}
% This material is based upon work supported by

\disclaimer{The views and conclusions contained in this document are those of the author and should not be interpreted as representing the official policies, either expressed or implied, of any sponsoring institution, the U.S.\ government or any other entity.}

\keywords{concurrency, bisimilarity, session types, proof construction, proof reduction, ordered logic, singleton logic}

\maketitle

\cleardoublepage
\begin{center}
\vspace*{2in}
\textit{For my parents and brother}
\end{center}

\makeatletter
\clearpage
\thispagestyle{empty}
\@ifundefined{@keywords}{}{\vspace*{\fill} \noindent{\bf Keywords:} \@keywords}
\cleardoublepage
\makeatother

\begin{abstract}
\normalsize
Concurrent systems are ubiquitous, but notoriously difficult to get right: subtle races and deadlocks can lurk even in the most extensively tested of systems.
In a quest to tame concurrency, researchers have successfully applied the principle of \emph{computation as deduction} to concurrency in two distinct ways: \emph{concurrency as proof reduction} and \emph{concurrency as proof construction}.
These two approaches to concurrency have complementary advantages, with the proof-construction approach excelling at global specification of a system's dynamics, while the proof-reduction approach is best suited to implementation of the processes that comprise the system.

This document explores the relationship between these two different proof-theoretic characterizations of concurrency.
To focus on the essential aspects of their relationship, the exploration is carried out in the context of concurrent systems that have chain topologies.
From a proof-construction perspective, chain topologies arise from ordered logic; from a proof-reduction perspective, they arise from \emph{singleton logic}, a variant of ordered logic that restricts sequents to have exactly one antecedent.

In this context, a rewriting framework is systematically derived from the ordered sequent calculus, and a \emph{message-passing} fragment of that rewriting framework is identified.
String rewriting specifications of concurrent systems can be \emph{choreographed} into this fragment, and the fragment supports a notion of bisimilarity.
Along the way, we also uncover a \emph{semi-axiomatic sequent calculus} for singleton logic, which blends a standard sequent calculus with axiomatic aspects of Hilbert systems, and we then establish a correspondence between semi-axiomatic proof normalization and asynchronous message-passing communication.
Ultimately, the message-passing processes can be faithfully embedded within the message-pass\-ing ordered rewriting framework in a bisimilar way.
Perhaps surprising is that, because the embedding is left-invertible, we can also identify fairly broad conditions under which local, process implementations can be \emph{extracted} from global, message-passing ordered rewriting specifications.
\end{abstract}

\begin{acknowledgments}
  \normalsize
  First, and foremost, the debt of gratitude that I owe Frank Pfenning cannot be overstated.
  Frank has been such a wonderful advisor and mentor to me throughout my time here at CMU, always patient and kind.
  Thank you, Frank, for inviting me on many a hike through the woods of proof theory and programming languages with you.

  I also wish to thank the members of my thesis committee -- Iliano Cervesato, Bob Harper, Andr\'{e} Platzer, Simon Gay, and Carsten Sch\"{u}rmann -- for being so generous with their time in talking with me and for their detailed feedback on my thesis document.  

  I am also grateful to my co-authors and the other people with whom I've had invaluable discussions over the years:
  Stephanie Balzer, Lu\'{\i}s Caires, Ankush Das, Anupam Datta, Farzaneh Derakhshan, Deepak Garg, Dilsun K\i rl\i\ Kaynar, Limin Jia, Andreia Mordido, Klaas Pruiksma, Rob Simmons, and Bernardo Toninho.

  Also, I would like to thank to Mark Stehlik for guiding me toward logic and programming languages.
  What still impresses me even now, all these years out of undergrad, is how he was able to recognize that it was just the field for me -- before I had any inkling of that myself.

  I am also thankful for everyone over the years who has helped me with scribing and typing my school work.

  Last, but most importantly, I want to thank my parents and my brother for all their love, care, and support through all these nearly 36 years -- not through only the good times, but most especially through the hard times.
  I love you.
\end{acknowledgments}

\oddsidemargin=\saveoddsidemargin
\evensidemargin=\saveevensidemargin

\tableofcontents
% \listoffigures
% \listoftables

\mainmatter
% \pagenumbering{arabic}

% \pagestyle{fancy}

\include{introduction}

\part{Preliminaries}\label{part:preliminaries}

\include{finite-automata}
\include{ordered-logic}

 \part[Concurrency as proof construction]{Concurrency as\\proof construction}\label{part:proof-construction}

\include{string-rewriting}
\include{ordered-rewriting-revised}
%% \include{ordered-rewriting}
% \include{move}
\include{choreographies}
% \include{garbage}
\include{ordered-bisimilarity}

\part[Concurrency as proof reduction]{Concurrency as\\proof reduction}\label{part:proof-reduction}

\include{singleton-logic}
\include{singleton-processes-exprs}
%% \include{singleton-processes}
%% \include{typed-bisim}

\part[Relationship between proof construction and reduction]{Relationship between\\proof construction\\and proof reduction}\label{part:comparison}

\include{from-procs-to-rewriting}
%% \include{session-typed-rewriting}
%% \include{transducer-isomorphism}
\include{conclusion}

\appendix
\include{ccs}

%% \include{misc}

\backmatter

\printbibliography

\end{document}
